---
title: "Test for temporal autocorrelation"
author: "Eleanor Jackson"
date: '`r format(Sys.time(), "%d %B, %Y")`'
always_allow_html: true
output: 
  github_document:
    keep_html: true
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.path = "figures/2021-06-23_test-temporal-autocorrelation/")
```

```{r}
library("tidyverse")
library("DHARMa")
library("glmmTMB")
library("broom.mixed")
library("patchwork")

fruit_traits <- readRDS(here::here("data", "clean", "fruit_traits.rds"))

set.seed(123)

```

## Fit models

```{r}
# round number of seeds to integers or else DHARMa will not recognise as binomial
fruit_traits$abscised_seeds <- round(fruit_traits$abscised_seeds)
fruit_traits$viable_seeds <- round(fruit_traits$viable_seeds)

# transform
fruit_traits %>%
    mutate(cofruit_cs = scale(cofruit),
    height_avg_cs = scale(height_avg),
    cvseed_cs = scale(cvseed),
    endocarp_investment_cs = scale(endocarp_investment),
    seed_dry_log = log(seed_dry),
    bcireproductive_log = log(bcireproductive)) %>%
    mutate(seed_dry_log_cs = scale(seed_dry_log),
    bcireproductive_log_cs = scale(bcireproductive_log),
    seedpred_pres = as.factor(seedpred_pres)) -> fruit_traits
```

```{r}
# select variables to loop over
vars <- c("height_avg_cs","cvseed_cs","cofruit_cs","endocarp_investment_cs",
        "seedpred_pres","seed_dry_log_cs","bcireproductive_log_cs")

vars <- setNames(vars, vars)

# loop model fitting over all the plant population attributes/traits

models <- lapply(setNames(vars, vars), function(var) {
	formula <- eval(parse(text=(paste("cbind(abscised_seeds, viable_seeds) ~", var,
					"+ (1|year) + (1|sp4)", sep = ""))))
	glmmTMB::glmmTMB(formula, data = fruit_traits, family = binomial(link = logit),
	                 na.action = "na.omit")
	}
)
```

## Test for temporal autocorrelation in the residuals

We need to get the residuals separately for each species in each model because testing for temporal autocorrelation requires unique time values.

I'm going to use the "normalised" residuals because they also take the random effect terms into account. See [here](https://stats.stackexchange.com/questions/80823/do-autocorrelated-residual-patterns-remain-even-in-models-with-appropriate-corre)

I'll use a Durbin-Watson test to look for autocorrelation.

```{r, warning=FALSE}
# function to get residuals
get_resids <- function(try_species, trait) {
  fruit_traits %>%
    drop_na(trait) %>%
    mutate(residuals = residuals(models[[trait]]), model = trait, type = "normalized") %>%
    dplyr::filter(sp4 == try_species)
}

keys <- expand.grid(unique(fruit_traits$sp4), vars) %>%
  rename(try_species = Var1, trait = Var2)

# loop over every combo of species and model
# not all species are in every model - so remove those empty dfs
pmap(keys, get_resids) %>% 
  keep(~ nrow(.x) > 0) -> resid_dfs

# perform a Durbin-Watson test on the residuals
test_autocor <- function(df){
  DHARMa::testTemporalAutocorrelation(df[["residuals"]], df[["year"]], plot = FALSE) -> out
  tibble(statistic = out[["statistic"]], p = out[["p.value"]],
         sp4 = df[["sp4"]][[1]], model = df[["model"]][[1]])
}

test_dfs <- lapply(resid_dfs, test_autocor)

autocor_vals <- bind_rows(test_dfs)

autocor_vals %>% 
  filter(p < 0.05 & statistic <1) %>%
  arrange(statistic) -> sig_pos_autocor_vals

autocor_vals %>% 
  filter(p < 0.05 & statistic >3) %>%
  arrange(statistic) -> sig_neg_autocor_vals

sig_pos_autocor_vals
sig_neg_autocor_vals

```

Out of `r nrow(autocor_vals)` values (one for each species in each model), `r nrow(sig_pos_autocor_vals)` are showing significant positive autocorrelation (`r signif(((nrow(sig_pos_autocor_vals)/nrow(autocor_vals))*100), digits = 3)`%) and `r nrow(sig_neg_autocor_vals)` are showing significant negative autocorrelation (`r signif(((nrow(sig_neg_autocor_vals)/nrow(autocor_vals))*100), digits = 3)`%). Looking at the tibbles above, it's interesting that some species show more temporal autocorrelation than others. They are also consistently temporally autocorrelated across models - which I think makes sense, as the same values of proportion abscised are used.

We can also look at the AutoCorrelation Function (ACF), which is calculated for each time lag.

```{r test-residual-autocorrelation}
# get autocorrelation values
test_autocor_acf <- function(df){
  acf(x = as.matrix(df[["residuals"]], df[["year"]]), plot = FALSE) -> out
  tibble(lag = out[["lag"]], acf = out[["acf"]],
         sp4 = df[["sp4"]][[1]], model = df[["model"]][[1]])
}

test_acf <- lapply(resid_dfs, test_autocor_acf)

acf_vals <- bind_rows(test_acf)

ggplot(acf_vals, aes(x=lag, y=acf, group = lag)) +
  geom_boxplot() +
  geom_hline(yintercept = 0, linetype="dashed") +
  geom_hline(yintercept = 0.5, linetype="dashed") +
  geom_hline(yintercept = -0.5, linetype="dashed") -> p1

p1
```

This doesn't look too bad - medians are close to zero, apart from at lag 2 where ACF is higher.

## Try fitting models with AR(1) covariance

I'm going to incorporate a  AR(1) covariance structure to see how it affects the autocorrelation and results.

```{r}
# fit with covariance
models_ar1 <- lapply(setNames(vars, vars), function(var) {
	formula <- eval(parse(text=(paste("cbind(abscised_seeds, viable_seeds) ~", var,
					"+ (1|sp4) + ar1(year + 0 | sp4)", sep = ""))))
	glmmTMB::glmmTMB(formula, data = fruit_traits, family = binomial(link = logit),
	                 na.action = "na.omit")
	}
)
```

## Test for autocorrelation in the new models

```{r, warning=FALSE}
get_resids_ar1 <- function(try_species, trait) {
  fruit_traits %>%
    drop_na(trait) %>%
    mutate(residuals = residuals(models_ar1[[trait]]), model = trait, type = "normalized") %>%
    dplyr::filter(sp4 == try_species)
}

pmap(keys, get_resids_ar1) %>%
  keep(~ nrow(.x) > 0) -> resid_dfs_ar1

# DW test
test_dfs_ar1 <- lapply(resid_dfs_ar1, test_autocor)

autocor_vals_ar1 <- bind_rows(test_dfs_ar1)

autocor_vals_ar1 %>% 
  filter(p < 0.05 & statistic <1) %>%
  arrange(statistic) -> sig_pos_autocor_vals_ar1

autocor_vals %>% 
  filter(p < 0.05 & statistic >3) %>%
  arrange(statistic) -> sig_neg_autocor_vals_ar1

sig_pos_autocor_vals_ar1
sig_neg_autocor_vals_ar1
```

Out of `r nrow(autocor_vals_ar1)` values (one for each species in each model), `r nrow(sig_pos_autocor_vals_ar1)` are showing significant positive autocorrelation. This is `r (nrow(sig_pos_autocor_vals)) -(nrow(sig_pos_autocor_vals_ar1))` less than the models without the AR(1) covariance structure. `r nrow(sig_neg_autocor_vals_ar1)` are showing significant negative correlation, the same as the models without the AR(1) covariance structure.

```{r ar1-covariance, out.width = "100%", warning=FALSE}
# autocorrelation values
test_autocor_acf <- function(df){
  acf(x = as.matrix(df[["residuals"]], df[["year"]]), plot = FALSE) -> out
  tibble(lag = out[["lag"]], acf = out[["acf"]],
         sp4 = df[["sp4"]][[1]], model = df[["model"]][[1]])
}

test_acf_ar1 <- lapply(resid_dfs_ar1, test_autocor_acf)

acf_vals_ar1 <- bind_rows(test_acf_ar1)

ggplot(acf_vals_ar1, aes(x=lag, y=acf, group = lag)) +
  geom_boxplot() +
  geom_hline(yintercept = 0, linetype="dashed") +
  geom_hline(yintercept = 0.5, linetype="dashed") +
  geom_hline(yintercept = -0.5, linetype="dashed") +
  ggtitle("with AR(1)") -> p2

p1 + p2
```

Some of the ACF values look better and some look worse. Lag 2 has been brought down a little.

## Compare model results

Note that I had to install a fix for `broom.mixed::tidy` to give confidence intervals for `glmmTMB` models. See [here](https://github.com/glmmTMB/glmmTMB/issues/401)

```{r compare-results}
#effects
res_anova <- purrr::map_dfr(models, effect="fixed", broom.mixed::tidy,
							conf.int = TRUE, .id = "vars")
results <- res_anova[!grepl("(Intercept)", res_anova$term),]

res_anova_ar1 <- purrr::map_dfr(models_ar1, effect="fixed", broom.mixed::tidy,
							conf.int = TRUE, .id = "vars")
results_ar1 <- res_anova_ar1[!grepl("(Intercept)", res_anova_ar1$term),]

results %>%
  mutate(model = "no_ar1") -> results

results_ar1 %>%
  mutate(model = "ar1") -> results_ar1

rbind(results, results_ar1) -> plot_dat

labs <- c(
            height_avg_cs = "Tree height",
            cvseed_cs = "Temporal crop size variation",
            cofruit_cs = "Temporal overlap in fruit\nproduction with other species",
            bcireproductive_log_cs = "Local abundance of conspecifics",
            endocarp_investment_cs = "Investment in mechanical\nseed defences",
            seed_dry_log_cs ="Seed mass",
            proportion_abscised = "Proportion of seeds abscised",
        	  seedpred_pres = "Presence of insect seed predator")

ggplot(plot_dat, aes(x = vars, y= estimate)) +
	geom_pointrange(aes(ymin = conf.low, ymax = conf.high),
        size = 0.5, fatten = 1)  +
	geom_hline(yintercept = 0, linetype = 2, size = 0.25)  +
    labs(y = "Estimate ± CI [95%]", x = "") +
    scale_x_discrete(labels = labs) +
    scale_y_continuous(limits = c(-1, 2.3)) +
    coord_flip() +
    facet_grid(~model) +
    theme_classic(base_size = 10)
```

Only a very small change in the results! Reassuring that it wouldn't change our conclusions if we decided to use the AR(1) covariance structure. However, I think that the autocorrelation we see in our model residuals is minimal and we don't need to adjust for it.
